MSE and dynamic state variable model with sigma function

### 2018-04-16 doc/MSE_MAR_18.tex
march

MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}. At the same time being simple enough to build and calibrate reliably with the available data. More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}. In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Kell2007, Dichmont2013}.

@@

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{Figures/MSE_6.eps} 
\caption{The conceptual framework and how it was mapped.}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{Figures/Areadynamics3.eps} 
\caption{The conceptual framework and how it was mapped.}
\label{fig:MSE}
\end{figure}

@@

To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}. A dynamic state variable model \cite{Houston1999, ClarkandMangel2000} was used to model location choice, extending the model structure in \cite{Batsleer2015}. In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations. The model had annual fines for exceeding landings or catch quota as in \cite{Alzorriz2018}). In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings or catch quotas at the end of the year:
\begin{equation}
\Phi (C_i, Q_i, F_i)= -\sum_i (max( 0, (C_i - Q_i))* F_i),
\end{equation}
where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel. These cumulative landings defined the state of the individual. $Q_i$ was the annual individual quota for landings or catches for the quota species. Individual quotas were not transferable. In order to reduce computation time, we set quota only for two species. $F_i$ was the fine per unit weight for exceeding individual landings or catch quotas was set to $3 x 10^3$ Euro day\textsuperscript{-1}. These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.

The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, s)= \Phi (C_i, Q_i, F_i)$. For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$ (A, B and including area 0: “staying in port”), and to keep or discard any combination of the age classes caught of the quota species. This discarding was defined by a matrix $d$. The size of that matrix was defined by the number of species under quota constraints and the number of age classes. Each element could take the value 0 (discard) or 1 (keep on board and land). The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
\begin{equation}
V (C_i, Q_i, F_i, s) = max_{p,d}( R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1])),
\end{equation}
where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependant over fish age (subsection \ref{sec2.2.1}). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table ~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices. The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

@@

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in ~\ref{eq:solve}. This was done for 700 individual vessels.

\begin{table}[!h]
\centering
\caption{Model and fishery conditioning.}
\label{t:fishery}
\resizebox{0.9\textwidth}{!}{%
\begin{tabular}{@{}lc lc lc @{}} 
\hline
\textbf{model}                              &               & \textbf{Fishery variable costs}                &      &\textbf{Fishery }&  \\ \hline
& & & \\%& & \\
number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  & q                 & 0.0005 \\ 
$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0    & effort (p,s)      & 1   \\
number of areas                             & 2             &crew share                                      & 35\% & CPUE & mean($pop*q*wts(yy-2:yy)$)\\
number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & CPUE dev& $0.08*CPUE$\\
intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
& & & \\\hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependant pricing}
\label{sec2.2.1}
Prices of fish have been assumed to be fixed over time but influenced by the size of the catch. Common knowledge and case studies indicate that size and weight of a fish could be important attributes when determining its economic value \cite{Zimmermann2011, Zimmermann2013}. Therefore, the resulting data of mean weights and average prices were used for further analysis. We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively. Thus, following \cite{Zimmermann2011}, the statistical models took the form:  
\begin{equation}
 pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
\end{equation}
where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes. $\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$. We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
 %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models. Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results. All models were fitted independently to each stock.
\begin{table}[!h]
\centering
\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}lcccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
                         & 1     & 2     & 3    & 4   \\ \hline
                      \\
species 1                &  778  & 1354   & 1780  & 2096  \\
species 2                &  778  & 1354   & 1780  & 2096  \\
\\
\hline 
\end{tabular}
}
\end{table}

@@

The three management strategies we considered differ by the formulae used to determine exploitation rates (HCR). These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan. For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.

\begin{table}[!h]
\centering
\caption{}
\label{scenarios}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c  c }
\hline
                                & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
& & &  \\
Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)                      &  $C = L$ ($D \neq 0$) \\
Selectivity ($l_{ratio}$)         &  $ \neq 1$                     & $=1$                                 & $=1$\\
Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\

&  & &  \\\hline
\end{tabular}
}
\end{table}

@@

\item 
\textbf{Scenario III: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding. However the manager could think that is fully enforced and assumed that only what is landed contribute to yield. This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality. Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.

@@

The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table ~\ref{t:fishery}) and the standing biomass:
\begin{equation}
 H_{a, y, s} =\frac
                % Nominator
                {C_{i (numbers)}(a, y, s, p)}
                % Denominator
                {C_{i (numbers)}(a, y, s, p) \times B_{i (numbers)}(a, y, s, p)},
\end{equation}

@@

\newpage
%\begin{landscape}
Scenario I:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand1.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand1.eps} 
\end{minipage}
\caption{}
\end{figure}
%\end{landscape}

@@

Scenario II:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO.eps} 
\end{minipage}
\caption{}
\end{figure}

Scenario III:

\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_nocompliance.eps}  
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\centering
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_nocompliance.eps} 
\end{minipage}
\caption{}
\label{fig:effort_lo}
\end{figure}

\subsection{Simulated catch decisions and location choice}
\label{sec3.1}

>>>>

MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}. At the same time being simple enough to build and calibrate reliably with the available data. More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}. In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Dichmont2013, Kell2007}.

@@

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{Figures/MSE.eps} 
\caption{The conceptual framework and how it was mapped.}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{Figures/Areadynamics.eps} 
\caption{The conceptual framework and how it was mapped.}
\label{fig:MSE}
\end{figure}

@@

To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}. A dynamic state variable model \cite{ClarkandMangel2000, Houston1999} was used to model location choice, extending the model structure in \cite{Batsleer2015}. In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations. The model had annual fines for exceeding landings or catch quota as in \cite{Alzorriz2018}). In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings or catch quotas at the end of the year:
\begin{equation}
\Phi (C_i, Q_i, F_i)= -\sum_i (max( 0, (C_i - Q_i))* F_i),
\end{equation}
where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel. These cumulative landings defined the state of the individual. $Q_i$ was the annual individual quota for landings or catches for the quota species. Individual quotas were not transferable. In order to reduce computation time, we set quota only for two species. $F_i$ was the fine per unit weight for exceeding individual landings or catch quotas was set to $3 x 10^3$ Euro day\textsuperscript{-1}. These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.

The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, s)= \Phi (C_i, Q_i, F_i)$. For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$ (A, B and including area 0: “staying in port”), and to keep or discard any combination of the age classes caught of the quota species. This discarding was defined by a matrix $d$. The size of that matrix was defined by the number of species under quota constraints and the number of age classes. Each element could take the value 0 (discard) or 1 (keep on board and land). The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
\begin{equation}
V (C_i, Q_i, F_i, s) = max_{p,d}( R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1])),
\end{equation}
where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependent over fish age (subsection \ref{sec2.2.1}). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices. The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

@@

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}. This was done for 700 individual vessels.

\begin{table}[!h]
\centering
\caption{Model and fishery conditioning.}
\label{t:fishery}
\resizebox{0.9\textwidth}{!}{%
\begin{tabular}{@{}lc lc lc @{}} 
\hline
\textbf{model}                              &               & \textbf{Fishery variable costs}                &      &\textbf{Fishery }&  \\ \hline
& & & \\%& & \\
number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 1240  & q                 & 0.0005 \\ 
$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0    & effort (p,s)      & 1   \\
number of areas                             & 2             &crew share                                      & 35\% & CPUE & mean($pop*q*wts(yy-2:yy)$)\\
number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & CPUE dev& $0.08*CPUE$\\
intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
& & & \\\hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependent pricing}
\label{sec2.2.1}
Prices of fish have been assumed to be fixed over time but influenced by the size of the catch. Common knowledge and case studies indicate that size and weight of a fish could be important attributes when determining its economic value \cite{Zimmermann2011, Zimmermann2013}. Therefore, the resulting data of mean weights and average prices were used for further analysis. We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively. Thus, following \cite{Zimmermann2011}, the statistical models took the form:  
\begin{equation}
 pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
\end{equation}
where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes. $\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$. We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
 %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models. Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results. All models were fitted independently to each stock.
\begin{table}[!h]
\centering
\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}llcccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
&                         & 1     & 2     & 3    & 4   \\ \hline
                      \\
Scenario I &              &       &        &       &     \\                       
&species 1                &  778  & 1354   & 1780  & 2096  \\
&species 2                &  778  & 1354   & 1780  & 2096  \\
                      \\
Scenario II &              &       &        &       &     \\ 
&species 1                &  778  & 1354   & 1780  & 2096  \\
&species 2                &  464  & 752    & 965   & 1123 \\
\\
\hline 
\end{tabular}
}
\end{table}

@@

The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR). These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan. For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.

\begin{table}[!h]
\centering
\caption{}
\label{scenarios}
\resizebox{0.8\textwidth}{!}{%
\begin{tabular}{l c c}
\hline
                                & \textbf{MP}            & \textbf{LO}                  \\ 
                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
& &   \\
Discarding                      & Allowed                        & Not allowed                 \\
Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\

&  &   \\\hline
\end{tabular}
}
\end{table}

@@

@@

The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table~\ref{t:fishery}) and the standing biomass:
\begin{equation}
 H_{a, y, s} =\frac
                % Nominator
                {C_{i (numbers)}(a, y, s, p)}
                % Denominator
                {C_{i (numbers)}(a, y, s, p) \times B_{i (numbers)}(a, y, s, p)},
\end{equation}

@@

We simulated the effects of implementing a Management Plan (MP) and the Landing Obligation (LO) without any exceptions or flexibility. To be able to analyse the results of this complex actions of implementing the Mp and the LO, the implementations should start simultaneously for all stocks and fleets. Under the MP, the Harvest Control Rule for the TACs is to advise $F_{MSY}$ unless the biomass falls below a trigger biomass. If the biomass falls below this value, the advised TAC is based on the biomass trends. To check for trends, the latest biomass levels are compared to previous ones. If there is a decrease of a 20\% or more, the TAC is reduced by 15\%. If the comparision shows an increase of more then 20 \% the TAC is increased by 15\%.

\newpage
%\begin{landscape}
Scenario I: prices of the two species are equal (a truly mixed fishery)
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioI.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioI.eps} 
\end{minipage}
\caption{}
\end{figure}
%\end{landscape}

@@

Scenario II: prices of the species 2 are 50\% of species 1 prices
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioII.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioII.eps} 
\end{minipage}
\caption{}
\end{figure}

\subsection{ catch decisions and location choice}
\label{sec3.1}
When discarding was allowed, but TACs changed by 15\% based on previous year but trying approximate to Fmsy, TACS showed a reduction and discards increased.

The implementation of LO caused an increase in TACs and revenues were higher than when discarding was allowed. Some fishing opportunities are always lost under the landing obligation; the final fishing mortalities are lower than the target fishing mortalities (at least for one species). However, this does not necessarily imply that the catches will be lower with landing obligation than without it. Catches will be reduced in the initial years of landing obligation implementation. However, at Spawning stock biomass increases with LO, the catches will increase too. Although the fishing opportunities are lost due to the choke effect, the biomass of some stocks might increase; after a few years, this extra abundance could be converted into more catches. This implies that are fleets that, in the mid-term, could benefit from the landing obligation.
